Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2021-09-09 manuscript.tex
Let's always write Sébastiens name correctly [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/4f0f379e0927ddbc18d27f425c700afd12fc7e42.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/4f0f379e0927ddbc18d27f425c700afd12fc7e42/checks
https://github.com/habi/acinar-analysis-manuscript/runs/1216506322

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...a1119e95c5134c8453ab9530b710fa95ce8a6770}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/a1119e95c5134c8453ab9530b710fa95ce8a6770/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/a1119e95c5134c8453ab9530b710fa95ce8a6770}{habi/acinar-analysis-manuscript@a1119e9}
on July 16, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(a1119e9)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(a1119e9)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

The total alveolar number for day 4 was found to be 0.64~±~0.34~million~alveoli, for day 10 1.23~±~0.39~million~alveoli, for day 21 2.65~±~0.57~million~alveoli, and for day 60 3.01~±~0.34~million~alveoli.
The total number of acini was estimated by dividing the mean parenchymal volume of the lungs {[}\protect\hyperlink{ref-wnl86DEM}{6}{]} by the stereologically estimated acinar volume.
The number of alveoli for each animal is then calculated by multiplying the average entrance ring count per animal with the estimated number of acini per animal.
To calculate this data we had to use data from three different studies performed by three different teams using two different strains of animals (i.e.~current data, Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]}, and Barre et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]}).
Therefore, we had to accept that error propagation may cause a significant blurring of our results.

@@

\begin{figure}
\hypertarget{fig:06}{%
\centering
\includegraphics{images/fig06.png}
\caption{Plot of the mean acinar volumes.
By dividing the parenchymal volumes of each lung estimated by Tschanz et al.~(Table 1, {[}\protect\hyperlink{ref-wnl86DEM}{6}{]}) by the mean number of acini estimated by Barre et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} we can estimate the mean acinar volume for each measured time point (blue plot).
We assessed the volumes of the acini by point counting (Cavalieri estimation, orange plot).
Left: original data.
Right: Our data, scaled by a linear correction factor.
Our volumes of the acini are on average 2.07 times lower but show a similar increase over the studied period.
While the estimation based on Tschanz et al.~is from the entire right middle lobe and on Barre et al.~globally for the whole organ, our estimation is solely based on the stereological analysis of the acini in the lower medial tip of the right lower lung lobe.}\label{fig:06}
}
\end{figure}

@@

Barrè et al.~have shown that the number of acini (e.g.~counted acinar entraces) remains constant during postnatal lung development from day 4 to 60 {[}\protect\hyperlink{ref-14OP85b2F}{10}{]}, with a mean of 5840~±~547~acini per lung (in another strain of animals, the data from the whole lungs is available for days 4 and 60).
Even though we did not aim to assess the number of acini in this manuscript, calculating this number helps to validate our assessment and helps us to collate the data that can be derived from our stereological analysis to prior results.

@@

\begin{figure}
\hypertarget{fig:07}{%
\centering
\includegraphics{images/fig07.png}
\caption{Plot of the estimated number of acini.
The blue plot shows our estimation, based on the division of the parenchymal volume by the average volume of the acini.
The orange plot is based on the data shown by Barre et al.~{[}\protect\hyperlink{ref-14OP85b2F}{10},\protect\hyperlink{ref-uFNlWogb}{11}{]}.}\label{fig:07}
}
\end{figure}

@@

However, relative numbers as reported in this study are more reliable.
This is especially the case when the comparison is done in groups of animals processed in parallel and counted by the same operator.
We assessed the same lungs as Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]}, but performed our assessment based on a different imaging method and with two different persons performing the stereological assessment at two different points in time.
Furthermore, we used the number of acini as determined by Barre et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} which represents a dataset obtained by a third team and which introduces the effect of error propagation.
This may explain the relative factor of 5.04 that we found as a difference.
Even if it was shown that the right lower lobe represents a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10},\protect\hyperlink{ref-I9TmP6IU}{45}{]}, we cannot rule out a contribution to the factor by regional differences in the lung which we can not account for in the present study.

@@

\leavevmode\hypertarget{ref-199ALtdJt}{}%
25. Flecknell P. Replacement, Reduction, Refinement. ALTEX. 2002;19: 73--78. Available: \url{https://www.altex.org/index.php/altex/article/view/1106}

@@

\leavevmode\hypertarget{ref-pQ6Wbz73}{}%
37. Kluyver T, Ragan-Kelley B, Pérez F, Granger B, Bussonnier M, Frederic J, et al. Jupyter Notebooks -- a publishing format for reproducible computational workflows. In: Loizides F, Scmidt B, editors. Positioning and Power in Academic Publishing: Players, Agents and Agendas. IOS Press; 2016. pp. 87--90. doi:\href{https://doi.org/10.3233/978-1-61499-649-1-87}{10.3233/978-1-61499-649-1-87}

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...4f0f379e0927ddbc18d27f425c700afd12fc7e42}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/4f0f379e0927ddbc18d27f425c700afd12fc7e42/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/4f0f379e0927ddbc18d27f425c700afd12fc7e42}{habi/acinar-analysis-manuscript@4f0f379}
on September 9, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(4f0f379)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(4f0f379)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

The total alveolar number for day 4 was found to be 0.64~±~0.34~million~alveoli, for day 10 1.23~±~0.39~million~alveoli, for day 21 2.65~±~0.57~million~alveoli, and for day 60 3.01~±~0.34~million~alveoli.
The total number of acini was estimated by dividing the mean parenchymal volume of the lungs {[}\protect\hyperlink{ref-wnl86DEM}{6}{]} by the stereologically estimated acinar volume.
The number of alveoli for each animal is then calculated by multiplying the average entrance ring count per animal with the estimated number of acini per animal.
To calculate this data we had to use data from three different studies performed by three different teams using two different strains of animals (i.e.~current data, Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]}, and Barré et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]}).
Therefore, we had to accept that error propagation may cause a significant blurring of our results.

@@

\begin{figure}
\hypertarget{fig:06}{%
\centering
\includegraphics{images/fig06.png}
\caption{Plot of the mean acinar volumes.
By dividing the parenchymal volumes of each lung estimated by Tschanz et al.~(Table 1, {[}\protect\hyperlink{ref-wnl86DEM}{6}{]}) by the mean number of acini estimated by Barré et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} we can estimate the mean acinar volume for each measured time point (blue plot).
We assessed the volumes of the acini by point counting (Cavalieri estimation, orange plot).
Left: original data.
Right: Our data, scaled by a linear correction factor.
Our volumes of the acini are on average 2.07 times lower but show a similar increase over the studied period.
While the estimation based on Tschanz et al.~is from the entire right middle lobe and on Barré et al.~globally for the whole organ, our estimation is solely based on the stereological analysis of the acini in the lower medial tip of the right lower lung lobe.}\label{fig:06}
}
\end{figure}

@@

Barré et al.~have shown that the number of acini (e.g.~counted acinar entrances) remains constant during postnatal lung development from day 4 to 60 {[}\protect\hyperlink{ref-14OP85b2F}{10}{]}, with a mean of 5840~±~547~acini per lung (in another strain of animals, the data from the whole lungs is available for days 4 and 60).
Even though we did not aim to assess the number of acini in this manuscript, calculating this number helps to validate our assessment and helps us to collate the data that can be derived from our stereological analysis to prior results.

@@

\begin{figure}
\hypertarget{fig:07}{%
\centering
\includegraphics{images/fig07.png}
\caption{Plot of the estimated number of acini.
The blue plot shows our estimation, based on the division of the parenchymal volume by the average volume of the acini.
The orange plot is based on the data shown by Barré et al.~{[}\protect\hyperlink{ref-14OP85b2F}{10},\protect\hyperlink{ref-uFNlWogb}{11}{]}.}\label{fig:07}
}
\end{figure}

@@

However, relative numbers as reported in this study are more reliable.
This is especially the case when the comparison is done in groups of animals processed in parallel and counted by the same operator.
We assessed the same lungs as Tschanz et al.~{[}\protect\hyperlink{ref-wnl86DEM}{6}{]}, but performed our assessment based on a different imaging method and with two different persons performing the stereological assessment at two different points in time.
Furthermore, we used the number of acini as determined by Barré et al.~{[}\protect\hyperlink{ref-uFNlWogb}{11}{]} which represents a dataset obtained by a third team and which introduces the effect of error propagation.
This may explain the relative factor of 5.04 that we found as a difference.
Even if it was shown that the right lower lobe represents a valid sample for the entire lung {[}\protect\hyperlink{ref-14OP85b2F}{10},\protect\hyperlink{ref-I9TmP6IU}{45}{]}, we cannot rule out a contribution to the factor by regional differences in the lung which we cannot account for in the present study.

@@

\leavevmode\hypertarget{ref-199ALtdJt}{}%
25. Flecknell P. Replacement, Reduction, Refinement. ALTEX - Alternatives to animal experimentation. 2002;19: 73--78. Available: \url{https://www.altex.org/index.php/altex/article/view/1106}

@@

\leavevmode\hypertarget{ref-pQ6Wbz73}{}%
37. Kluyver T, Ragan-Kelley B, Pérez F, Granger B, Bussonnier M, Frederic J, et al. Jupyter Notebooks -- a publishing format for reproducible computational workflows. In: Loizides F, Scmidt B, editors. IOS Press; 2016. pp. 87--90. doi:\href{https://doi.org/10.3233/978-1-61499-649-1-87}{10.3233/978-1-61499-649-1-87}
